----
title: "plots of annual temperature summaries, USGS"
output:
  html_document:
    toc: true
    theme: united
---- 

# Basic Analysis
how are Annual Mean (and Min/Max) stream and air temperatures changing over time? and, what is the relationship between stream and air temperature change?

```{r}

library(dplyr)
library(feather)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(plotly)
library(Kendall)
source("./src/helpers.R")

# load in USGS-Daymet data
daily <- read_feather("./data/usgs_daymet_all.feather")
annual <- read_feather("./data/summary/annual_temp_z.feather")

# sites with > 40 years
site_n <- annual %>%
  group_by(site_code) %>%
  summarise(n = n()) %>%
  filter(n > 40)

# for adding to df for plotting
pmap <- function(v) {
  pvals <- c()
  for(x in v) {
    if(x <= 0.05) {
      pvals <- c(pvals, 'significant')
    } else {
      pvals <- c(pvals, 'non-significant')
    }
  }
  return(pvals)
}

slopemap <- function(v) {
  slopes <- c()
  for(x in v) {
    if(x > 1) {
      slopes <- c(slopes, 'positive')
    } else if(x < 1){
      slopes <- c(slopes, 'negative')
    } else {
      slopes <- c(slopes, 'zero')
    }
  }
  return(slopes)
}

```

## Annual Trend Analsysis
let's look at plots of the Mean, Min, and Max Annual temps.

We can also look at the same data transformed to Z-score (a metric of the distance of a particular point to it's dataset mean, values between -x and x, with 0 representing the perfect mean, and -x and x representing x standard deviations below and above the mean respectively), in this way normalizing all site data to itself to facilitate mroe direct comparison of temperature chamge/deviation between sites.

### mean

lets filter the data to only site with > 40 year records, before 2022, and, lets calculate the p-values of regressions fit to the relationship between a site's meanimum temperature and the year

```{r}
# keep only records > certain period of years
annual.mean <- annual %>%
  filter(year < 2022) %>%
  filter(site_code %in% site_n$site_code)

# set infinite values to NA
annual.mean$mean_wtr_air_ratio[is.infinite(annual.mean$mean_wtr_air_ratio)] <- NA
# remove NA
annual.mean <- annual.mean %>%
  select(site_code, year, mean_wtr_temp, mean_mean_air_temp, mean_wtr_air_ratio) %>%
  tidyr::drop_na()


# singificane and slope testing
# Determeane p-values of regression
p.vals = sapply(unique(annual.mean$site_code), function(i) {
  coef(summary(lm(year ~ mean_wtr_temp, data=annual.mean[annual.mean$site_code==i, ])))[2,4]
})

slope.vals = sapply(unique(annual.mean$site_code), function(i) {
  coef(summary(lm(year ~ mean_wtr_temp, data=annual.mean[annual.mean$site_code==i, ])))[2,1]
})

# do same for ratio
ratio.p.vals = sapply(unique(annual.mean$site_code), function(i) {
  coef(summary(lm(year ~ mean_wtr_air_ratio, data=annual.mean[annual.mean$site_code==i, ])))[2,4]
})

ratio.slope.vals = sapply(unique(annual.mean$site_code), function(i) {
  coef(summary(lm(year ~ mean_wtr_air_ratio, data=annual.mean[annual.mean$site_code==i, ])))[2,1]
})

annual.mean <- annual.mean %>%
  mutate(
    pval = p.vals[site_code],
    pcat = pmap(pval),
    slope = slope.vals[site_code],
    slopecat = slopemap(slope),
    # ratio
    ratio.pval = ratio.p.vals[site_code],
    ratio.pcat = pmap(ratio.pval),
    ratio.slope = ratio.slope.vals[site_code],
    ratio.slopecat = slopemap(ratio.slope)
  )



wtr_mean_plot <- ggplotly(
      ggplot(annual.mean, aes(x=year, y=mean_wtr_temp, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.mean,
        aes(year, mean_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Mean Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.mean$year)[seq(1, 50, 10)])
)
  
print(wtr_mean_plot)

air_mean_plot <- ggplotly(
      ggplot(annual.mean, aes(x=year, y=mean_mean_air_temp, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.mean,
        aes(year, mean_mean_air_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Mean Air Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.mean$year)[seq(1, 50, 10)])
)
  
print(air_mean_plot)

combo_mean_plot <- ggplotly(
      ggplot(annual.mean, aes(x=year, group=site_code)) +
      ## geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.mean,
        aes(year, mean_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      geom_smooth(
        data=annual.mean,
        aes(year, y=mean_mean_air_temp, colour=site_code),
        method='lm',
        se = FALSE,
        linetype = "dashed"
      ) +
      xlab("Year") +
      ggtitle("Annual Mean Air and Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      ## facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.mean$year)[seq(1, 50, 10)])
)
  
print(combo_mean_plot)

# water:air temp ratio
ratio_mean_plot <- ggplotly(
      ggplot(annual.mean, aes(x=year, y=mean_wtr_air_ratio, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.mean,
        aes(year, mean_wtr_air_ratio, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Mean of Water:Air Temperature Ratio Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ ratio.slopecat + ratio.pcat) +
      scale_x_discrete(breaks = unique(annual.mean$year)[seq(1, 50, 10)])
)
  
print(ratio_mean_plot)

# real TS
ts_ratio_mean_plot <- ggplotly(
      ggplot(annual.mean, aes(x=year, y = mean_wtr_air_ratio, group=site_code)) +
      geom_line(aes(color = site_code)) +
      xlab("Year") +
      ggtitle("Annual Mean Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ ratio.slopecat + ratio.pcat) +
      scale_x_discrete(breaks = unique(annual.mean$year)[seq(1, 50, 10)])
)

print(ts_ratio_mean_plot)
```

### min

```{r}
# keep only records > certain period of years
annual.min <- annual %>%
  filter(year < 2022) %>%
  filter(site_code %in% site_n$site_code) %>%
  mutate(
    wtr_air_min_ratio = min_wtr_temp/min_tmin..deg.c.
  )

# singificane and slope testing
# Determine p-values of regression
p.vals = sapply(unique(annual.min$site_code), function(i) {
  coef(summary(lm(year ~ min_wtr_temp, data=annual.min[annual.min$site_code==i, ])))[2,4]
})

slope.vals = sapply(unique(annual.min$site_code), function(i) {
  coef(summary(lm(year ~ min_wtr_temp, data=annual.min[annual.min$site_code==i, ])))[2,1]
})

# for ratio
ratio.p.vals = sapply(unique(annual.min$site_code), function(i) {
  coef(summary(lm(year ~ wtr_air_min_ratio, data=annual.min[annual.min$site_code==i, ])))[2,4]
})

ratio.slope.vals = sapply(unique(annual.min$site_code), function(i) {
  coef(summary(lm(year ~ wtr_air_min_ratio, data=annual.min[annual.min$site_code==i, ])))[2,1]
})

```
    
```{r}
annual.min <- annual.min %>%
  mutate(
    pval = p.vals[site_code],
    pcat = pmap(pval),
    slope = slope.vals[site_code],
    slopecat = slopemap(slope),
    # ratio
    ratio.pval = ratio.p.vals[site_code],
    ratio.pcat = pmap(ratio.pval),
    ratio.slope = ratio.slope.vals[site_code],
    ratio.slopecat = slopemap(ratio.slope)
  )

wtr_min_plot <- ggplotly(
      ggplot(annual.min, aes(x=year, y=min_wtr_temp, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.min,
        aes(year, min_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Min Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.min$year)[seq(1, 50, 10)])
)
  
print(wtr_min_plot)

air_min_plot <- ggplotly(
      ggplot(annual.min, aes(x=year, y=min_tmin..deg.c., group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.min,
        aes(year, min_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Min Air Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.min$year)[seq(1, 50, 10)])
)
  
print(air_min_plot)

combo_min_plot <- ggplotly(
      ggplot(annual.min, aes(x=year, group=site_code)) +
      ## geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.min,
        aes(year, min_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      geom_smooth(
        data=annual.min,
        aes(year, y=min_tmin..deg.c., colour=site_code),
        method='lm',
        se = FALSE,
        linetype = "dashed"
      ) +
      xlab("Year") +
      ggtitle("Annual Min Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.min$year)[seq(1, 50, 10)])
)
  
print(combo_min_plot)

# water:air temp ratio
ratio_min_plot <- ggplotly(
      ggplot(annual.min, aes(x=year, y=wtr_air_min_ratio, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.min,
        aes(year, wtr_air_min_ratio, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Mean of Water:Air Minimum Temperature Ratios Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ ratio.slopecat + ratio.pcat) +
      scale_x_discrete(breaks = unique(annual.min$year)[seq(1, 50, 10)])
)
  
print(ratio_min_plot)

# real TS
ts_ratio_min_plot <- ggplotly(
      ggplot(annual.min, aes(x=year, y = wtr_air_min_ratio, group=site_code)) +
      geom_line(aes(color = site_code)) +
      xlab("Year") +
      ggtitle("Annual Min Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ ratio.slopecat + ratio.pcat) +
      scale_x_discrete(breaks = unique(annual.min$year)[seq(1, 50, 10)])
)

print(ts_ratio_min_plot)
```

### max
let's look at a plot of the Z-score (a metric of the distance of a particular point to it's dataset mean, values between -2 and 2, with 0 representing the perfect mean, and -2 and 2 representing two standard deviations below and above respectively)

```{r}
# keep only records > certain period of years
annual.max <- annual %>%
  filter(year < 2022) %>%
  filter(site_code %in% site_n$site_code) %>%
  mutate(
    wtr_air_max_ratio = max_wtr_temp/max_tmax..deg.c.
  )

# singificane and slope testing
# Determaxe p-values of regression
p.vals = sapply(unique(annual.max$site_code), function(i) {
  coef(summary(lm(year ~ max_wtr_temp, data=annual.max[annual.max$site_code==i, ])))[2,4]
})

slope.vals = sapply(unique(annual.max$site_code), function(i) {
  coef(summary(lm(year ~ max_wtr_temp, data=annual.max[annual.max$site_code==i, ])))[2,1]
})

# for ratio
ratio.p.vals = sapply(unique(annual.max$site_code), function(i) {
  coef(summary(lm(year ~ wtr_air_max_ratio, data=annual.max[annual.max$site_code==i, ])))[2,4]
})

ratio.slope.vals = sapply(unique(annual.max$site_code), function(i) {
  coef(summary(lm(year ~ wtr_air_max_ratio, data=annual.max[annual.max$site_code==i, ])))[2,1]
})


annual.max <- annual.max %>%
  mutate(
    pval = p.vals[site_code],
    pcat = pmap(pval),
    slope = slope.vals[site_code],
    slopecat = slopemap(slope),
    # ratio
    ratio.pval = ratio.p.vals[site_code],
    ratio.pcat = pmap(ratio.pval),
    ratio.slope = ratio.slope.vals[site_code],
    ratio.slopecat = slopemap(ratio.slope)
  )

wtr_max_plot <- ggplotly(
      ggplot(annual.max, aes(x=year, y=max_wtr_temp, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.max,
        aes(year, max_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Max Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.max$year)[seq(1, 50, 10)])
)
  
print(wtr_max_plot)

air_max_plot <- ggplotly(
      ggplot(annual.max, aes(x=year, y=max_tmax..deg.c., group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.max,
        aes(year, max_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Max Air Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.max$year)[seq(1, 50, 10)])
)
  
print(air_max_plot)

combo_max_plot <- ggplotly(
      ggplot(annual.max, aes(x=year, group=site_code)) +
      ## geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.max,
        aes(year, max_wtr_temp, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      geom_smooth(
        data=annual.max,
        aes(year, y=max_tmax..deg.c., colour=site_code),
        method='lm',
        se = FALSE,
        linetype = "dashed"
      ) +
      xlab("Year") +
      ggtitle("Annual Max Water and Air Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ slopecat + pcat) +
      scale_x_discrete(breaks = unique(annual.max$year)[seq(1, 50, 10)])
)
  
print(combo_max_plot)

# water:air temp ratio
ratio_max_plot <- ggplotly(
      ggplot(annual.max, aes(x=year, y=wtr_air_max_ratio, group=site_code)) +
      geom_point(aes(color = site_code)) +
      geom_smooth(
        data=annual.max,
        aes(year, wtr_air_max_ratio, colour=site_code),
        method='lm',
        se = FALSE
      ) +
      xlab("Year") +
      ggtitle("Annual Max Water:Air Temperature Ratios Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ ratio.slopecat + ratio.pcat) +
      scale_x_discrete(breaks = unique(annual.max$year)[seq(1, 50, 10)])
)
  
print(ratio_max_plot)

# real TS
ts_ratio_max_plot <- ggplotly(
      ggplot(annual.max, aes(x=year, y = wtr_air_max_ratio, group=site_code)) +
      geom_line(aes(color = site_code)) +
      xlab("Year") +
      ggtitle("Annual Max Water Temperatures Across USGS Sites") +
      theme_minimal() +
      theme(text = element_text(size = 30)) +
      facet_wrap(~ ratio.slopecat + ratio.pcat) +
      scale_x_discrete(breaks = unique(annual.max$year)[seq(1, 50, 10)])
)

print(ts_ratio_max_plot)
```
