# Seasonal Mann Kendall Daily Data
```{r}
library(dplyr)
library(feather)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(plotly)
library(Kendall)
library(trend)
source("./src/helpers.R")

# load in USGS-Daymet data
daily <- read_feather("./data/usgs_daymet_all.feather")
annual <- read_feather("./data/summary/annual_temp_z.feather")

# Mann Kendall & Sen's Slope, Stream Temp
annual.wtr <- annual %>%
  select(year, site_code, mean_wtr_temp)

sites <- unique(annual.wtr$site_code)

mk.data <- data.frame(
  site_code = c(''),
  mk_tau = c(''),
  mk_sl = c(''),
  mk_S = c(''),
  mk_D = c(''),
  sens_slope = c(''),
  sens_conf = c(''),
  sens_p = c(''),
  sens_z = c(''),
  sens_n = c('')
)

for(i in 1:length(sites)) {
  
    annual.ts <- annual.wtr[annual.wtr$site_code == sites[i],]  %>%
        arrange(year) %>%
        tidyr::drop_na() %>%
        select(mean_wtr_temp) %>%
        ts(frequency = 1,
           start = min(annual.wtr$year),
           end = max(annual.wtr$year),
           )

    annual.mk <- MannKendall(annual.ts)
    annual.sens <- sens.slope(annual.ts)
  
    tryCatch(
      expr = {
        # Mann Kendall
        mk.data[i,]$site_code <- sites[i]
        mk.data[i,]$mk_tau <- annual.mk[[1]][1]
        mk.data[i,]$mk_sl <- annual.mk[[2]][1]
        mk.data[i,]$mk_S <- annual.mk[[3]][1]
        mk.data[i,]$mk_D <- annual.mk[[4]][1]

        # Sen's Slope
        mk.data[i,]$sens_slope <- annual.sens$estimates[[1]]
        mk.data[i,]$sens_conf <- annual.sens$conf.int[[1]]
        mk.data[i,]$sens_p <- annual.sens$p.value[[1]]
        mk.data[i,]$sens_z <- annual.sens$statistic[[1]]
        mk.data[i,]$sens_n <- annual.sens$parameter[[1]]
      },
      error = function(e) {
        print(paste("ERROR:", sites[i]))
    }
  ) 
}


write.csv(mk.data, "mannkendall_sens_annual_wtr.csv")

# look at sig trends
mk.annual.sig <- mk.data %>%
  filter(mk_sl < .05) %>%
  filter(sens_p < .05)

plot(mk.annual.sig$sens_slope, mk.annual.sig$mk_tau)
```

repeat, for air tmeperature data

```{r}

# Mann Kendall & Sen's Slope, Air Temp
annual.air <- annual %>%
  select(year, site_code, mean_mean_air_temp)

sites <- unique(annual.air$site_code)

mk.data.air <- data.frame(
  site_code = c(''),
  mk_tau = c(''),
  mk_sl = c(''),
  mk_S = c(''),
  mk_D = c(''),
  sens_slope = c(''),
  sens_conf = c(''),
  sens_p = c(''),
  sens_z = c(''),
  sens_n = c('')
)

for(i in 1:length(sites)) {
  
    annual.ts <- annual.air[annual.air$site_code == sites[i],]  %>%
        arrange(year) %>%
        tidyr::drop_na() %>%
        select(mean_mean_air_temp) %>%
        ts(frequency = 1,
           start = min(annual.air$year),
           end = max(annual.air$year),
           )

    annual.mk <- MannKendall(annual.ts)
    annual.sens <- sens.slope(annual.ts)
  
    tryCatch(
      expr = {
        # Mann Kendall
        mk.data.air[i,]$site_code <- sites[i]
        mk.data.air[i,]$mk_tau <- annual.mk[[1]][1]
        mk.data.air[i,]$mk_sl <- annual.mk[[2]][1]
        mk.data.air[i,]$mk_S <- annual.mk[[3]][1]
        mk.data.air[i,]$mk_D <- annual.mk[[4]][1]

        # Sen's Slope
        mk.data.air[i,]$sens_slope <- annual.sens$estimates[[1]]
        mk.data.air[i,]$sens_conf <- annual.sens$conf.int[[1]]
        mk.data.air[i,]$sens_p <- annual.sens$p.value[[1]]
        mk.data.air[i,]$sens_z <- annual.sens$statistic[[1]]
        mk.data.air[i,]$sens_n <- annual.sens$parameter[[1]]
      },
      error = function(e) {
        print(paste("ERROR:", sites[i]))
    }
  ) 
}


write.csv(mk.data.air, "mannkendall_sens_annual_air.csv")

# look at sig trends
mk.annual.sig.air <- mk.data.air %>%
  filter(mk_sl < .05) %>%
  filter(sens_p < .05)

plot(mk.annual.sig.air$sens_slope, mk.annual.sig.air$mk_tau)

```


# Daily, SMK
daily.smk <- daily %>%
  select(date, site_code, wtr_temp)

# Create time series that matches your data
# x= start date, "mm/dd/yyyy"
# x= end date, "mm/dd/yyyy"

timer <- function(x,y){
  data.frame(
    date = as.Date(seq(as.POSIXlt(x, format="%Y-%m-%d"), as.POSIXlt(y, format="%Y-%m-%d"), by = 'day'))
    )
}


# Seasonal Mann Kendall
library(mice)
library(trend)

daily.ts <- merge(ts, daily.smk, by="date")
daily.ts.f <- daily.ts %>%
  tidyr::drop_na()
library(xts)


test.mk <- daily.ts.f %>%
  filter(site_code == "01481000") %>%
  select(-site_code) %>%
  arrange(date)

start.mk <- min(test.mk$date)
end.mk <- max(test.mk$date)
ts <- timer(start, end)

test <- ts(test.mk$wtr_temp, frequency = 365, start = c(1971, 1))
test.smk <- SeasonalMannKendall(test)
test.sn <- sens.slope(test)

# test
x <- AutoSMK(daily.smk, ts, "wtr_temp", "01481000", start, end)

# run SMK for each site
sites <- unique(daily$site_code)

mk.data <- data.frame(
  site_code = c(''),
  start_date = c(''),
  end_date = c(''),
  smk_tau = c(''),
  smk_sl = c(''),
  smk_S = c(''),
  smk_D = c(''),
  sen_z = c(''),
  sen_n = c(''),
  sen_p = c(''),
  sen_slope = c(''),
  sen_conf = c('') 
)

for(i in 1:length(sites)) {
  
    wtr <- daily[daily$site_code == sites[i],] %>%
        select(date, wtr_temp) %>%
        arrange(date) %>%
        tidyr::drop_na()
  
    tryCatch(
      expr = {
        mk <- SeasonalMannKendall(wtr$wtr_temp)
        mk.data[i,]$site_code <- sites[i]
        mk.data[i,]$tau <- mk[[1]][1]
        mk.data[i,]$sl <- mk[[2]][1]
        mk.data[i,]$S <- mk[[3]][1]
        mk.data[i,]$D <- mk[[4]][1]

        mk <- MannKendall(wtr$wtr_temp)
        mk.data[i,]$site_code <- sites[i]
        mk.data[i,]$tau <- mk[[1]][1]
        mk.data[i,]$sl <- mk[[2]][1]
        mk.data[i,]$S <- mk[[3]][1]
        mk.data[i,]$D <- mk[[4]][1]
      },
      error = function(e) {
        print(paste("ERROR:", sites[i]))
    }
  ) 
}

```
